---
title: "Files_Count"
author: "Edwin King'ori"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---



File_Counts Visual Analysis

The 'File_Count' table in the Data Summary sheet provides a complete set of data required for a comprehensive visual analysis to be conducted. For this reason, various visual graphs were generated to find the relationships of the given variables.   

Loading the necessary Library Packages
```{r}
library(readxl)
library(dplyr)
library(tidyverse)
library(janitor)
library(skimr)
library(ggplot2)
library(tidyr)
library(readr)
library(reshape2)
library(scales)
```

Importing the data-set

```{r}
files_count_df <- read_xlsx("Files_counts_summary.xlsx")
head(files_count_df)
```
Inspecting the Data-Set
```{r}
summary(files_count_df)
str(files_count_df)
```
The data is clean and credible for the next visualization phase
However, since the data is wide, it has to be transformed to a long format for it to be compatible with the desired plot.

In this case:
    Reshaping the data to long format for it to display the legend in the visualization.
    
```{r}
files_count_long_df <- files_count_df %>%
  pivot_longer(cols = c(Gifts, Promotions, Donors),
               names_to = "Category",
               values_to = "Count")
```

#1 Generating a horizontal bar plot to compare the gifts, donor and  promotions across all sectors

```{r}

ggplot(files_count_long_df, aes(x = Sector,y = Count, fill = Category))+
  geom_bar(stat = "identity", position = "dodge")+
  scale_y_continuous(labels = scales::comma) + 
  scale_fill_manual(values = c("Gifts" = "blue", "Donors" = "red", "Promotions" = "green")) +
  theme_minimal()+
  labs(title = "Comparing the gifts, donor and  promotions across all sectors",
       x = "Sector", y = "Count")+
  theme(axis.text.x = element_text(angle = 90, hjust = 0.1))+ 
  
  coord_flip()
```

2 Visualizing the Correlation of Gifts, Donors, Promotions and Sustainable Donor Count by Sector using a Stacked bar plot

```{r}
# transforming the data from a wide data to a long data using the melt function
files_count_long <- melt(files_count_df, id.vars = "Sector", measure.vars = c("Gifts", "Donors", "Promotions", "SustainerDonorCount"))
```


```{r}
ggplot(files_count_long, aes(x = Sector, y = value, fill = variable)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Breakdown of Gifts, Donors, Promotions, and Sustainer Donor Count by Sector", x = "Sector", y = "Count")+
  scale_y_continuous(labels = comma)+
  scale_fill_manual(values = c("Gifts" = "blue", "Donors" = "red", "Promotions" = "green", "SustainerDonorCount" = "black"))

```

3 Visualizing the  Relationship between Gifts and Donors

```{r}
ggplot(files_count_df, aes(x = Gifts, y = Donors, color = Sector)) +
  geom_point(size = 4) +
  theme_minimal() +
  labs(title = "Relationship between Gifts and Donors", x = "Gifts", y = "Donors")+
  scale_y_continuous(labels = comma)+
  scale_x_continuous(labels = comma)+
  theme(axis.text.x = element_text(angle =90 , hjust = 0.1))

```
4 Visualizing the  Relationship between Gifts and Promotions

```{r}
ggplot(files_count_df, aes(x = Gifts, y = Promotions, color = Sector)) +
  geom_point(size = 4) +
  theme_minimal() +
  labs(title = "Relationship between Gifts and Promotions", x = "Gifts", y = "Promotions")+
  scale_y_continuous(labels = comma)+
  scale_x_continuous(labels = comma)+
  theme(axis.text.x = element_text(angle =90 , hjust = 0.1))

```
5. Visualizing the  Relationship between  Donors and Promotions

```{r}

ggplot(files_count_df, aes(x = Donors, y = Promotions, color = Sector)) +
  geom_point(size = 4) +
  theme_minimal() +
  labs(title = "Relationship between Donors and Promotions", x = "Donors", y = "Promotions")+
  scale_y_continuous(labels = comma)+
  scale_x_continuous(labels = comma)+
  theme(axis.text.x = element_text(angle =90 , hjust = 0.1))

```

6. Heatmap: Correlation of matrix

```{r}

file_matrix <- cor(files_count_df[, c("Gifts", "Donors", "Promotions", "SustainerDonorCount")])
ggplot(melt(file_matrix), aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-1, 1), space = "Lab", name="Correlation") +
  theme_minimal() +
  labs(title = "Heatmap of Correlation Matrix", x = "", y = "")

```
Visual Translation

In all the plots, the variables Gifts, Donors and Promotions exhibit strong positive relationship, which is indicated by the higher values in these categories correlating with each other. However, there are noticeable differences of counts across the different sectors. For instance, the veterans receive a higher rate of promotions compared to sectors like health, which receive a higher number of gifts. . Despite the huge count differences between the variables, the intense red coloring exhibited in the heatmap confirms a positive correlation of the three variables. The sustainerDonorCount variable  shows a moderate correlation with other variables but a much stronger relationship with the Donors variable. 


